The Dynamics of Anthropomorphic Effects in Human-Robot Interaction

### 2014-03-25 paper.tex
new changes in anthropomorphic design section ... almost done

Most work that studies how far human users treat robots as social agents, builds on the 
famous work of \cite{reeves_media_1996} in HCI. In the so-called \textit{Media 
Equation}, the authors describe that people mindlessly react in a social way to their computer 
and that they tend to even attribute human (social) characteristics to it -- despite 
the fact that a computer does not have human-like characteristics. Apart from 
computers, anthropomorphism has also been observed with digital / virtual characters 
\citep{reeves_media_1996,nass_anthropocentrism_1995}, and artificial embodied agents, 
such as robots. In short, one can say that with robots, and human-like robots in 
particular, anthropomorphism is a commonly observed phenomenon. 
But what are the motivations which underlie the efforts of enhancing anthropomorphism 
through specific anthropomorphic design in robots? Why are anthropomorphic forms so 
distinct and at the same time controversial in robotics? There is no one single answer 
to these questions.

@@

On one hand, developers and designers have discovered that human-like robots may better 
match the affordances that our to humans adapted world brings along. Also, up to a 
certain point, anthropomorphic form as it increases anthropomorphism,
appears to be generally beneficial for human-robot relations, especially with socially 
interactive robots \cite{fong_survey_2003}. Consequently, the domain of (social) 
robotics tries to exploit this fact by enhancing anthropomorphism through 
anthropomorphic robots: Applying anthropomorphism or zoomorphism changes the 
way how users try to understand and make sense of a system by projecting everyday 
\hl{expectations} of human and animalistic life onto it. \cite{schmitz_concepts_2011}.

In general, it is suggested that increasing the human-likeness of a robot (up to a 
certain threshold) will lead to 
more perceived \textit{affinity} from the human side \citep{mori_uncanny_1970} (see 
Figure \ref{fig:uncanny_valley}). Humans seem to readily accept and prefer the 
anthropomorphic form of a robot, when compared to purely mechanical robots. A 
possible explanation for this is that 
humans tend to apply the same communication references to anthropomorphic systems 
that they would use with other humans, and this is what they are familiar with and feel comfortable about. This theoretical concept is also supported 
by psychological determinants of 
perceiving human-likeness in something non-human: \textit{"the more similar in 
appearance, the more people are likely to use themselves as a source of induction 
and anthropomorphize these nonhuman agents"} \citep{epley_seeing_2007}. (We will come back to this later, in Section \ref{sec:explanations}).

Studies have further shown that perceived similarity between the user and the robot influences the intensity of the participants' attributions towards the robot. Thus,the anthropomorphic design of a robot impacts the user's perception of the robot, and how far the user is likely to attribute and experience qualities such as intelligence, fun, or intentions in the robot.

"[...] the experiment demonstrates that human beings implicitly attribute human-like qualities, such as mental states, to nonhuman agents. This finding was evident on a neuro-physiological level as well as behavioral level." \cite{hegel_understanding_2008}\fxwarning{What to do with this quote? 'neuro' level maybe interesting?}

Various 
empirical studies have shown that robots with human-like forms can have a positive 
effect, as they enhance social 
responses from humans which in turn can positively impact on acceptance 
\citep{duffy_anthropomorphism_2003,goetz_cooperation_2002,venkatesh_theoretical_2000}. 
People responded more positively to an artifact that displayed human-like 
behavioral characteristics (emotions, facial expression) in contrast to a purely 
functional design 
\citep{eyssel_anthropomorphic_2010,krach_can_2008,reeves_media_1996,riek_how_2009}. 
Overall, these results suggest that human-like features in robots enhance 
anthropomorphism, and in turn create a social view toward the robot.

Anthropomorphism represents just one of many examples of induction whereby "people reason about an unknown stimulus based on a better-known representation of a related stimulus" \cite{epley_when_2008}, in this case reasoning about a non-human agent based on representation of the self or other humans.

@@

Anthropomorphism and similar phenomenons can also be understood as a thoughtful process of induction whereby \textit{``people reason about an unknown stimulus based on a better-known representation of a related stimulus"}~\citep{epley_when_2008}. In the case of anthropomorphism this means a person's reasoning about a non-human agent based on the representation of the self or other humans. The general assumption here is that all other agents (and entities) but oneself are \textit{``unknown''} (because we can only be our self). A central construct of this perspective is the so-called \textit{mental model} that a person has (and builds) of the agents she is reasoning about and interacting with. The specific mental model we have about an artifact / agent basically explains how this artifact / agent works the way it does. But how can we explain the phenomenon that a person uses a human-inspired mental model to explain the actions of a non-human (unknown) agent? Well, what we know and understand best is ourselves as human beings (the mental model that we have of our self and other humans). So by applying a similar mental model (or parts of it) to the unknown agent, we seek to be able to explain this unknown agent.
This view is consistent with the \emph{familiarity thesis}~\citep{hegel_understanding_2008} which claims that we understand the world based upon a mental model of the world that we are most familiar with. Also, the mental model (human-centered perspective) is supported by \cite{epley_seeing_2007} \textit{Three Factor Theory} which applies a social psychology view, and will be presented later in this section.

@@

%\paragraph*{The mental model theory in human-robot interaction\\ \\}

>>>>

Most work that studies how far human users treat robots as social agents, builds on the 
famous work of \cite{reeves_media_1996} in HCI. In the so-called \textit{Media 
Equation}, the authors describe that people mindlessly react in a social way to their computer 
and that they tend to even attribute human (social) characteristics to it -- despite 
the fact that a computer does not have human-like characteristics. Apart from 
computers, anthropomorphism has also been observed with digital / virtual characters 
\citep{von_der_putten_it_2010}, and artificial embodied agents, 
such as robots \citep{rosenthal-vonderputten_experimental_2013}. 
In short, one can say that with robots, and human-like robots in 
particular, anthropomorphism is a commonly observed phenomenon. 
But what are the motivations which underlie the efforts of enhancing anthropomorphism 
through specific anthropomorphic design in robots? Why are anthropomorphic forms so 
distinct and at the same time controversial in robotics? There is no one single answer 
to these questions.

@@

On one hand, developers and designers have discovered that human-like robots may better 
match the affordances that this to humans adapted world brings along. 
In general, it is suggested that increasing the human-likeness of a robot (up to a 
certain threshold) will lead to 
more perceived \textit{affinity} from the human side \citep{mori_uncanny_1970} (see 
Figure \ref{fig:uncanny_valley}).
This further increases anthropomorphism, and as such
appears to be beneficial for human-robot relations, especially with socially 
interactive robots \cite{fong_survey_2003}. Consequently, the domain of (social) 
robotics tries to exploit this by using
anthropomorphic forms in robots.
Humans seem to readily accept and prefer the 
anthropomorphic form of a robot (up to a certain point), when compared to purely mechanical 
robots. A possible explanation for this is that perceiving a system as
human-like (or also pet-like) compared to machine-like changes the 
way how users try to understand and make sense of this system. The reason for this is 
because humans tend to project everyday 
\hl{expectations} of human and animalistic life onto artificial agents that use 
anthropomorphic (or zoomorphic) forms \cite{schmitz_concepts_2011}. 
Both cognitive science 
and social psychology can provide explanations for this process of projection. The basic idea 
is that \textit{"the more similar in 
appearance, the more people are likely to use themselves as a source of induction 
and anthropomorphize these nonhuman agents"} \citep{epley_seeing_2007}. Studies have further 
shown that perceived similarity between the user and the robot influences the intensity of 
the participants' attributions towards the robot. Thus,the anthropomorphic design of a robot 
impacts the user's perception of the robot, and how far the user is likely to attribute and 
experience qualities such as intelligence, fun, or intentions in the robot.
(We will come back to this more thoughtful reaction to anthropomorphic forms, in more detail 
later, in Section \ref{sec:human-centered}).

In short, the central point for the design of sociable robots is that 
humans tend to apply the same communication references to anthropomorphic robots 
that they would use with other humans, which in turn enhances feeling familiar and 
comfortable when interacting with the otherwise unknown system.
This positive effect has been shown in various 
empirical studies. It has been found that robots with human-like forms can enhance social 
responses from humans which in turn can positively impact on acceptance 
\citep{duffy_anthropomorphism_2003,goetz_cooperation_2002,venkatesh_theoretical_2000}. 
People responded more positively to an artifact that displayed human-like 
behavioral characteristics (emotions, facial expression) in contrast to a purely 
functional design 
\citep{eyssel_anthropomorphic_2010,krach_can_2008,reeves_media_1996,riek_how_2009}. 
Overall, these results suggest that human-like features in robots enhance 
anthropomorphism, and in turn create a social view toward the robot.

%Anthropomorphism represents just one of many examples of induction whereby "people reason about an unknown stimulus based on a better-known representation of a related stimulus" \cite{epley_when_2008}, in this case reasoning about a non-human agent based on representation of the self or other humans.

@@

Anthropomorphism and similar phenomenons can also be understood as a thoughtful process of induction whereby \textit{``people reason about an unknown stimulus based on a better-known representation of a related stimulus"}~\citep{epley_when_2008}. In the case of anthropomorphism this means a person's reasoning about a non-human agent based on the representation of the self or other humans. The general assumption here is that all other agents (and entities) but oneself are \textit{``unknown''} (because we can only be our self). A central construct of this perspective is the so-called \textit{mental model} that a person has (and builds) of the agents she is reasoning about and interacting with. The specific mental model we have about an artifact / agent basically explains how this artifact / agent works the way it does. But how can we explain the phenomenon that a person uses a human-inspired mental model to explain the actions of a non-human (unknown) agent? Well, what we know and understand best is ourselves as human beings (the mental model that we have of our self and other humans). So by applying a similar mental model (or parts of it) to the unknown agent, we seek to be able to explain this unknown agent.
This view is consistent with the \emph{familiarity thesis}~\citep{hegel_understanding_2008} which claims that we understand the world based upon a mental model of the world that we are most familiar with. The authors found that human beings implicitly attribute human-like qualities, such as mental states, to nonhuman agents. This finding was evident on a neuro-physiological level as well as behavioral level. \footnote{\hl{Ha! This says that anthropomorphism can also occur on a behavioral level, yes! :)}} Also, the mental model (human-centered perspective) is supported by \cite{epley_seeing_2007} \textit{Three Factor Theory} which applies a social psychology view, and will be presented later in this section.

@@
